Comparison and Evaluation of Popular Syntax Highlighting Themes in Terms of Color Correlated Temperature Values

### 2022-04-26 doc/manuscript.tex
.

\documentclass{article}
\usepackage{hyperref}
\usepackage{ifthen, array, booktabs}
\usepackage{csvsimple}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{fancyvrb}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\usepackage{comment}
\usepackage{svg}
\svgpath{{./img/}}
\usepackage{pgfplots, pgfplotstable}
\pgfplotsset{compat=1.17}

@@

The first step was to choose a theme. Themes are generally classified into two theme groups according to the use of dark/light background colors in the user interface: dark/light themes.  Theme selection is dependent on editor selection. The VS Code programming editor was chosen for our research\footnote{Rationales behind all these selections were explained in further sections.}. Developers use a variety of programming languages to create the code for their projects. Because analyzing the influence of the chosen theme on all codes would be impractical, operations were conducted out on a sample source code. For this purpose, a C programming language source code taken from the Git source code, which is a fairly modern code base, was chosen. The HTML output for the colorized version of the code with the chosen theme, as well as a code representation based on it, was created. Different fonts are preferred by developers. The influence of different fonts on the calculation was not investigated in this study. The BDF transformation and subsequently the font representation have been created for the selected font. The McCammy equation has been used to determine the CCT using code and font representation.

@@

@@

\subsubsection{Calculation of font character fill rate}

\begin{figure}[htb]
  \centering
  \begin{BVerbatim}
#include "cache.h"
#include "config.h"
···
static void std_output(struct grep_opt *opt, const void *buf, size_t size)
{
        fwrite(buf, size, 1, stdout);
}
  \end{BVerbatim}
  \caption{Sample code without coloring}
  \label{code:string}
\end{figure}

\begin{figure}[h]
  \centering
  \includesvg[inkscapelatex=false, width=370pt]{code.svg}
  \caption{Colored code}
  \label{fig:code}
\end{figure}

In syntax coloring, each syntax group (e.g., an identifier, a keyword) in the source code is displayed with the font
selected in color set in the theme. To calculate the contribution of a character in the group to the color temperature,
it is necessary to calculate the character bit density from the bitmap of the character in the selected font. To explain
the method followed in making this calculation, for example, the definition of BDF given in Figure
\ref{code:letter_a_BDF} with the Consolas font of the character '\texttt{a}` is discussed.

In this definition, each line starting with \texttt{BITMAP} corresponds to a line in the bit matrix of the corresponding
character. A bitmap is obtained when bit patterns reported in hexadecimal are converted to binary (Figure
\ref{fig:letter_a_bitmap}). Based on this bitmap, the appearance of character '\texttt{a}' is shown in Figure \ref{fig:letter_a}.

\begin{figure}[htbp]
  \caption{BDF definition of the Consolas font for the character '\texttt{a}'}
  \label{code:letter_a_BDF}

  \centering
  \begin{BVerbatim}
    STARTCHAR 0061
    ENCODING 97
    SWIDTH 525 0
    DWIDTH 7 0
    BBX 5 7 1 0
    BITMAP
    70
    88
    08
    78
    88
    88
    F8
    ENDCHAR
  \end{BVerbatim}
\end{figure}

\begin{figure}[htbp]
  \caption{Bit pattern in Consolas font of character '\texttt{a}'}
  \label{fig:letter_a_bitmap}

  \centering
  \begin{BVerbatim}
    0 1 1 1 0 0 0 0
    1 0 0 0 1 0 0 0
    0 0 0 0 1 0 0 0
    0 1 1 1 1 0 0 0
    1 0 0 0 1 0 0 0
    1 0 0 0 1 0 0 0
    1 1 1 1 1 0 0 0
  \end{BVerbatim}
\end{figure}

\begin{figure}[H]
  \centering
  \includesvg[inkscapelatex=false, width=100pt]{line.svg}
  \caption{Line of colored character boxes}
  \label{fig:line}
\end{figure}

\begin{figure}[htbp]
    \centering
    \includesvg[inkscapelatex=false, width=150pt]{charbox.svg}
    \caption{Consolas TTF of '\texttt{a}' character with line height}
    \label{fig:charbox}
\end{figure}

\colorlet{color0}{yellow!40}
\colorlet{color1}{red}

\begin{figure}[htbp]
  \centering
  \caption{Consolas BDF of '\texttt{a}' character with line height}
  \label{fig:letter_a}

  \begin{tikzpicture}[%
    scale=0.35,
    box/.style={rectangle,draw=black,thin, minimum size=1cm},
    every node/.style={transform shape}
  ]%

  \foreach \x in {0, 1, ..., 8}{
    \foreach \y in {0, 1, ...,21}{
      \def\bit{0} \pgfmathsetmacro{\bit}{\pixelmap_a[mod(\y,22)][mod(\x,9)]}
      \node[box, fill=color\bit] at (\x,\y){};
    }
  }
  \end{tikzpicture}
\end{figure}

Renklendirilmiş kayna kodun ortalama CCT değeri Eşitlik \ref{eq:cct_average} ile elde edilmiştir.

Bu ifadede görülen $CCT(charbox_{i})$ kaynak kod dizgisindeki $i$'nci karakter kutusunun ortalama CCT değerine olan CCT
katkısını ifade etmekte olup Eşitlik \ref{eq:cct_charbox} ile tanımlanmıştır.

@@

Bu ifadede $i$ renklendirilmiş kaynak kod dizgisindeki $i$'nci karakter kutusuna karşı gelmek üzere:

@@

>>>>

\documentclass{article}
\usepackage{hyperref}
\usepackage{ifthen, array, booktabs}
\usepackage{csvsimple}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{fancyvrb}
\usepackage{amsmath}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\usepackage{comment}
\usepackage{subcaption}
\usepackage{svg}
\svgpath{{./img/}}
\usepackage{pgfplots, pgfplotstable}
\pgfplotsset{compat=1.17}

@@

The first step was to choose a theme. Themes are generally classified into two theme groups according to the use of
dark/light background colors in the user interface: dark/light themes.  Theme selection is dependent on editor
selection. The VS Code programming editor was chosen for our research. Developers use a variety of programming languages
to create the code for their projects. Because analyzing the influence of the chosen theme on all codes would be
impractical, operations were conducted out on a sample source code. For this purpose, a C programming language source
code taken from the Git source code, which is a fairly modern code base, was chosen. The HTML output for the colorized
version of the code with the chosen theme, as well as a code representation based on it, was created. Different fonts
are preferred by developers. The influence of different fonts on the calculation was not investigated in this study. The
BDF transformation and subsequently the font representation have been created for the selected font. The McCammy
equation has been used to determine the CCT using code and font representation.

@@

\begin{figure}[htb]
  \centering
  \begin{BVerbatim}
#include "cache.h"
#include "config.h"
···
static void std_output(struct grep_opt *opt, const void *buf, size_t size)
{
        fwrite(buf, size, 1, stdout);
}
  \end{BVerbatim}
  \caption{Sample code without coloring}
  \label{fig:plain_code}
\end{figure}

\begin{figure}[h]
  \centering
  \includesvg[inkscapelatex=false, width=370pt]{code.svg}
  \caption{Colored code}
  \label{fig:colored_code}
\end{figure}

@@

\subsubsection{Calculation of CCT contribution of a character box}

In syntax coloring, each syntax group (e.g. an identifier, a keyword) in the source code is displayed with the font
glyph selected in color set in the theme. Bölüm \ref{sec:material-methods} başlangıcında da belirtildiği gibi kaynak kod
karakter dizisindeki her bir karakter içinde bulunduğu satırın yüksekliğine sahip bir karakter kutusu oluşturmakta ve
buna göre renklendirilmiş kaynak kod içeriğinin tamamı bu kutulardan oluşan bir karakter kutusu dizisi olarak
yorumlanabilmektedir. Şekil \ref{fig:colored_code}'de verilen renklendirilmiş kod örneği dikkate alındığında bu karakter
kutusu dizisinde "\texttt{static}" anahtar kelimesine ilişkin bir bölüm Şekil \ref{fig:line_of_charboxes}'de
verilmiştir.

\begin{figure}[H]
  \centering
  \includesvg[inkscapelatex=false, width=100pt]{line.svg}
  \caption{Line of colored character boxes}
  \label{fig:line_of_charboxes}
\end{figure}

Ortalama CCCT değeri iki adımda hesaplanmaktadır.

\begin{enumerate}
  \item Her bir karakter kutusunun ortalama CCT değerinin hesaplanması için kutuda bulunan font glyph'ın ön ve arka plan
    renklerinin CCT katkısı bu renklerin kutu içinde kapladığı piksel alanı oranları dikkate alınarak hesaplanmalıdır.
    Şekil \ref{fig:line_of_charboxes}'deki örnekte ön plan rengi yeşil, arka plan rengi beyaz olarak verilmiştir.

  \item Her bir karakter kutusunun içeriğe olan CCT katkısı, karakter kutusunun içerikteki sıklığı dikkate alınarak
    hesaplanmalıdır.  Örneğin Şekil \ref{fig:line_of_charboxes}'de verilen 6 karakterlik içerikte \texttt{a} karakterine
    ait karakter kutusu 2 kez, \texttt{s} karakterine ait kutu ise 1 kez geçmektedir.
\end{enumerate}

Birinci adımda yapılan hesaplamayı örneklemek için \texttt{a} karakteri için Şekil \ref{fig:charbox_TTF}'de verilen
karakter kutusu dikkate alınsın.  Bu karakterin BDF font tanımı ve bu tanımda yer alan bit eşlemi Şekil
\ref{fig:letter_a_BDF}'deki gibidir.

\begin{figure}[htbp]
  \begin{subfigure}{0.48\textwidth}
    \centering
    \begin{BVerbatim}
      STARTCHAR 0061
      ENCODING 97
      SWIDTH 525 0
      DWIDTH 7 0
      BBX 5 7 1 0
      BITMAP
      70
      88
      08
      78
      88
      88
      F8
      ENDCHAR
    \end{BVerbatim}
    \caption{BDF definition}
    \label{code:letter_a_BDF_definitiion}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
    \centering
    \begin{BVerbatim}
    0 1 1 1 0 0 0 0
    1 0 0 0 1 0 0 0
    0 0 0 0 1 0 0 0
    0 1 1 1 1 0 0 0
    1 0 0 0 1 0 0 0
    1 0 0 0 1 0 0 0
    1 1 1 1 1 0 0 0
    \end{BVerbatim}
    \caption{Bit map}
    \label{fig:letter_a_BDF_bitmap}
  \end{subfigure}

  \caption{Consolas font of character '\texttt{a}' in BDF}
  \label{fig:letter_a_BDF}
\end{figure}

Renklendirilmiş kaynak kodun HTML temsilinde bildirilen \texttt{line-height} değeri ve Şekil
\ref{fig:letter_a_BDF_bitmap}'deki bit eşlemi birlikte dikkate alınacak olursa \texttt{a} karakteri için Şekil
\ref{fig:charbox_BDF}'deki karakter kutusu elde edilir.  Bu karakter kutusunun içerik dizisinde $i$'nci karakter kutusu
olduğu kabul edilirse bu kutuya ait $CCT(charbox_{i})$ ile gösterilen ortalama CCT katkısı Eşitlik \ref{eq:cct_charbox}
ile tanımlanmıştır.

@@

Bu ifadede:

@@

İkinci adımda, her bir karakter kutusu için ilk adımda hesaplanan CCT değerlerinin ortalaması alınmalıdır. Karakter
kutusu dizisinin $N$ boyutlu olduğu kabul edilecek olursa tüm içeriğin ortalama CCT değeri, yani CCCT değeri Eşitlik
\ref{eq:cct_average} ile elde edilir.

\colorlet{color0}{black!10}
\colorlet{color1}{black!40!green}

\begin{figure}[htbp]
  \begin{subfigure}{0.48\textwidth}
    \centering
    \includesvg[inkscapelatex=false, width=145pt]{charbox.svg}

    \caption{TTF}
    \label{fig:charbox_TTF}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
    \centering
    \begin{tikzpicture}[%
      scale=0.35,
      box/.style={rectangle,draw=black,thin, minimum size=1cm},
      every node/.style={transform shape}
    ]%

    \foreach \x in {0, 1, ..., 8}{
      \foreach \y in {0, 1, ...,21}{
        \def\bit{0} \pgfmathsetmacro{\bit}{\pixelmap_a[mod(\y,22)][mod(\x,9)]}
        \node[box, fill=color\bit] at (\x,\y){};
      }
    }
    \end{tikzpicture}

    \caption{BDF}
    \label{fig:charbox_BDF}
  \end{subfigure}

  \caption{Character box of '\texttt{a}' with line height in Consolas fontface}
  \label{fig:charbox}
\end{figure}
